A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-08-15 tidyindex.qmd
migrate changes from .text to qmd

-   $x$ is the variable in $\mathbb{R}^p$,

-   $s$ represents the geographic locations in the space $D_s \subseteq \mathbb{R}^m$. This includes a collection of countries of interest and paired longitude and latitude coordinates of locations, and

-   $t$ denotes the temporal order in $D_t \subseteq \mathbb{R}^n$. Sometimes, it can be time breaks into other temporal components, such as year, month, quarter, and season.

@@

| No. | Module                  | Input                                                       | Operation                                                                                                                                          | Output                                                                  |
|---------------|---------------|---------------|---------------|---------------|
| 1   | Temporal processing     | $x(t)$                                                      | $f_{\Psi}[x(.)]$                                                                                                                                   | $x^*(t^\prime)$ where $t\prime \in D_{t^\prime}$                        |
| 2   | Spatial processing      | $x(s)$                                                      | $g_{\Theta}[x(.)]$                                                                                                                                 | $x^\#(s^\prime)$ where $s^\prime \in D_{s^\prime}$                      |
| 3   | Variable transformation | $x(s; t)$                                                   | $T(x(s;t))$                                                                                                                                        | $u(s;t)$ ????                                                           |
| 4   | Scaling                 | $x(s; t)$                                                   | $[x(s;t) - \alpha]/\gamma$                                                                                                                         | $v(s;t)$ where bbbbb                                                    |
| 5   | Dimension reduction     | $\mathbf{x}(s;t)$ where $\mathbf{x} \subseteq \mathbb{R}^p$ | aaaaa                                                                                                                                              | $\mathbf{y}(s;t)$ where $\mathbf{y} \subseteq \mathbb{R}^d$ and $d < p$ |
| 6   | Distribution fit        | $x(s; t)$                                                   | $F_\mathbf{\theta}(x(s,t))$ where $F_{\mathbf{\theta}}(.)$ is the cumulated distribution function with distribution parameter(s) $\mathbf{\theta}$ | $p(s,t)$ where $p(.)$ denotes the probability value between 0 and 1     |
| 7   | Normalising             | $x(s; t)$                                                   | $\Phi^{-1}[x(s, t)]$                                                                                                                               | $z(s; t)$ where $z \sim N(.,.)$                                         |
| 8   | Benchmarking            | $x(s; t)$                                                   |                                                                                                                                                    | $b(s;t)$                                                                |
| 9   | Simplification          | $x(s; t)$                                                   | ccccc                                                                                                                                              | $A(s;t) \in \{a_1, a_2, \cdots, a_j\}$                                  |

>>>>

-   $x^i(s, t)$ represents a variable of interest for example precipitation,$j = 1, \cdots, p$ and

-   $s$ represents the geographic locations in the space $D_s \subseteq \mathbb{R}^m$. Examples of geographic locations include a collection of countries, longitude and latitude coordinates or regions of interest and,

-   $t$ denotes the temporal order in $D_t \subseteq \mathbb{R}^n$. For instance, time measurements could be recorded hourly, yearly, monthly, quaterly or by season..

@@

| No. | Module                  | Input                                                       | Operation                                                                                                                                          | Output                                                                  |
|------------|------------|------------|--------------------------|------------|
| 1   | Temporal processing     | $x(t)$                                                      | $f_{\Psi}[x(.)]$                                                                                                                                   | $x^*(t^\prime)$ where $t\prime \in D_{t^\prime}$                        |
| 2   | Spatial processing      | $x(s)$                                                      | $g_{\Theta}[x(.)]$                                                                                                                                 | $x^\#(s^\prime)$ where $s^\prime \in D_{s^\prime}$                      |
| 3   | Variable transformation | $x(s; t)$                                                   | $T(x(s;t))$                                                                                                                                        | $u(s;t)$ ????                                                           |
| 4   | Scaling                 | $x(s; t)$                                                   | $[x(s;t) - \alpha]/\gamma$                                                                                                                         | $v(s;t)$ where bbbbb                                                    |
| 5   | Dimension reduction     | $\mathbf{x}(s;t)$ where $\mathbf{x} \subseteq \mathbb{R}^p$ | aaaaa                                                                                                                                              | $\mathbf{y}(s;t)$ where $\mathbf{y} \subseteq \mathbb{R}^d$ and $d < p$ |
| 6   | Distribution fit        | $x(s; t)$                                                   | $F_\mathbf{\theta}(x(s,t))$ where $F_{\mathbf{\theta}}(.)$ is the cumulated distribution function with distribution parameter(s) $\mathbf{\theta}$ | $p(s,t)$ where $p(.)$ denotes the probability value between 0 and 1     |
| 7   | Normalising             | $x(s; t)$                                                   | $\Phi^{-1}[x(s, t)]$                                                                                                                               | $z(s; t)$ where $z \sim N(.,.)$                                         |
| 8   | Benchmarking            | $x(s; t)$                                                   |                                                                                                                                                    | $b(s;t)$                                                                |
| 9   | Simplification          | $x(s; t)$                                                   | ccccc                                                                                                                                              | $A(s;t) \in \{a_1, a_2, \cdots, a_j\}$                                  |

### 2023-08-15 tidyindex.tex
migrate changes from .text to qmd

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\item
  \(x\) is the variable in \(\mathbb{R}^p\),
\item
  \(s\) represents the geographic locations in the space
  \(D_s \subseteq \mathbb{R}^m\). This includes a collection of
  countries of interest and paired longitude and latitude coordinates of
  locations, and
\item
  \(t\) denotes the temporal order in \(D_t \subseteq \mathbb{R}^n\).
  Sometimes, it can be time breaks into other temporal components, such
  as year, month, quarter, and season.
\end{itemize}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
\caption{An notation overview of the input, operation, and output of
each pipeline module.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endhead
1 & Temporal processing & \(x(t)\) & \(f_{\Psi}[x(.)]\) &
\(x^*(t^\prime)\) where \(t\prime \in D_{t^\prime}\) \\
2 & Spatial processing & \(x(s)\) & \(g_{\Theta}[x(.)]\) &
\(x^\#(s^\prime)\) where \(s^\prime \in D_{s^\prime}\) \\
3 & Variable transformation & \(x(s; t)\) & \(T(x(s;t))\) & \(u(s;t)\)
???? \\
4 & Scaling & \(x(s; t)\) & \([x(s;t) - \alpha]/\gamma\) & \(v(s;t)\)
where bbbbb \\
5 & Dimension reduction & \(\mathbf{x}(s;t)\) where
\(\mathbf{x} \subseteq \mathbb{R}^p\) & aaaaa & \(\mathbf{y}(s;t)\)
where \(\mathbf{y} \subseteq \mathbb{R}^d\) and \(d < p\) \\
6 & Distribution fit & \(x(s; t)\) & \(F_\mathbf{\theta}(x(s,t))\) where
\(F_{\mathbf{\theta}}(.)\) is the cumulated distribution function with
distribution parameter(s) \(\mathbf{\theta}\) & \(p(s,t)\) where
\(p(.)\) denotes the probability value between 0 and 1 \\
7 & Normalising & \(x(s; t)\) & \(\Phi^{-1}[x(s, t)]\) & \(z(s; t)\)
where \(z \sim N(.,.)\) \\
8 & Benchmarking & \(x(s; t)\) & & \(b(s;t)\) \\
9 & Simplification & \(x(s; t)\) & ccccc &
\(A(s;t) \in \{a_1, a_2, \cdots, a_j\}\) \\
\bottomrule()
\end{longtable}

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\item
  \(x^i(s, t)\) represents a variable of interest for example
  precipitation,\(j = 1, \cdots, p\) and
\item
  \(s\) represents the geographic locations in the space
  \(D_s \subseteq \mathbb{R}^m\). Examples of geographic locations
  include a collection of countries, longitude and latitude coordinates
  or regions of interest and,
\item
  \(t\) denotes the temporal order in \(D_t \subseteq \mathbb{R}^n\).
  For instance, time measurements could be recorded hourly, yearly,
  monthly, quaterly or by season..
\end{itemize}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1622}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1622}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1622}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3514}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1622}}@{}}
\caption{An notation overview of the input, operation, and output of
each pipeline module.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endhead
1 & Temporal processing & \(x(t)\) & \(f_{\Psi}[x(.)]\) &
\(x^*(t^\prime)\) where \(t\prime \in D_{t^\prime}\) \\
2 & Spatial processing & \(x(s)\) & \(g_{\Theta}[x(.)]\) &
\(x^\#(s^\prime)\) where \(s^\prime \in D_{s^\prime}\) \\
3 & Variable transformation & \(x(s; t)\) & \(T(x(s;t))\) & \(u(s;t)\)
???? \\
4 & Scaling & \(x(s; t)\) & \([x(s;t) - \alpha]/\gamma\) & \(v(s;t)\)
where bbbbb \\
5 & Dimension reduction & \(\mathbf{x}(s;t)\) where
\(\mathbf{x} \subseteq \mathbb{R}^p\) & aaaaa & \(\mathbf{y}(s;t)\)
where \(\mathbf{y} \subseteq \mathbb{R}^d\) and \(d < p\) \\
6 & Distribution fit & \(x(s; t)\) & \(F_\mathbf{\theta}(x(s,t))\) where
\(F_{\mathbf{\theta}}(.)\) is the cumulated distribution function with
distribution parameter(s) \(\mathbf{\theta}\) & \(p(s,t)\) where
\(p(.)\) denotes the probability value between 0 and 1 \\
7 & Normalising & \(x(s; t)\) & \(\Phi^{-1}[x(s, t)]\) & \(z(s; t)\)
where \(z \sim N(.,.)\) \\
8 & Benchmarking & \(x(s; t)\) & & \(b(s;t)\) \\
9 & Simplification & \(x(s; t)\) & ccccc &
\(A(s;t) \in \{a_1, a_2, \cdots, a_j\}\) \\
\bottomrule()
\end{longtable}
